---
title: "Project2"
output: html_document
---

# Monte Carlo Feature Selection

Identify the set of relevant features that would be used for classification

```{r}

install.packages("rJava")
install.packages("rmcfs")
library(rJava)
library(rmcfs)

# Load data
df <- read.csv("data/clean/positions.csv")

# Drop column ID
positions <- subset(df, select = -id )

# Perform monte carlo feature selection (random reducts)
result <- mcfs(class ~ ., positions, projections = 1500, projectionSize = 0.1, splits = 5, splitSetSize = 500, cutoffPermutations = 6, threadsNumber = 8)

```

```{r}

# Render results
plot(result, type = "distances")

# Fetch results
result2 <- result$RI[1:result$cutoff_value,]

# Keep only relevant features
selected_positions <- positions[, colnames(positions) %in% c(result2$attribute, "class"), drop = FALSE]

```

# R.ROSETTA Basic Unit Test

Create and train a rule-based model with the selected features 

```{r}



```

# Optimize Model Parmeters

Refine the rule-based model based on empirical results

```{r}


```

# Clustering Model Rules

Show clustering with heatmap

```{r}


```

# Visualization

Visualize rules (boxplots, VisuNet, VisuArc)

```{r}


```

