---
title: "3.3 - Parameters Grid Search"
output: html_document
---

# Cell 1: Setup and Data

```{r}
set.seed(435)
library(R.ROSETTA)

selected_positions <- read.csv("data/clean/selected_positions.csv")

cat("Data:", nrow(selected_positions), "rows x", ncol(selected_positions), "cols\n")
cat("Classes:", paste(names(table(selected_positions$class)), table(selected_positions$class), sep="=", collapse=", "), "\n")
```

# ============================================================
# GRID SEARCH 1: Johnson BRT / BRTprec / Fraction (64 configs)
# ============================================================

# Cell 2: Build Grid 1

```{r}
fraction_vals <- c(0.99, 0.95, 0.9, 0.8, 0.7, 0.5, 0.25, 0.1)
brtprec_vals  <- c(1, 0.9, 0.8, 0.6, 0.5, 0.25, 0.1)
discern_vals  <- c("Full", "Object")

grid1_brt_false <- expand.grid(
    BRT                   = FALSE,
    BRTprec               = 0,
    Fraction              = fraction_vals,
    reducerDiscernibility = discern_vals,
    stringsAsFactors = FALSE
)

grid1_brt_true <- expand.grid(
    BRT                   = TRUE,
    BRTprec               = brtprec_vals,
    Fraction              = fraction_vals,
    reducerDiscernibility = discern_vals,
    stringsAsFactors = FALSE
)

grid1 <- rbind(grid1_brt_false, grid1_brt_true)
cat("Grid 1 configs:", nrow(grid1), "\n")
```

# Cell 3: Run Grid 1

```{r}
results1 <- data.frame(
    BRT                   = logical(),
    BRTprec               = numeric(),
    Fraction              = numeric(),
    reducerDiscernibility = character(),
    accuracyMean          = numeric(),
    nRules                = integer(),
    nSigRules             = integer(),
    stringsAsFactors = FALSE
)

for (i in 1:nrow(grid1)) {
    cfg <- grid1[i, ]
    cat(sprintf("Grid1 %d/%d — BRT=%s BRTprec=%.2f Fraction=%.2f Discern=%s\n",
                i, nrow(grid1), cfg$BRT, cfg$BRTprec, cfg$Fraction, cfg$reducerDiscernibility))

    jp <- list(
        Modulo      = TRUE,
        BRT         = cfg$BRT,
        BRTprec     = cfg$BRTprec,
        Precompute  = FALSE,
        Approximate = TRUE,
        Fraction    = cfg$Fraction
    )

    ros <- tryCatch(
        rosetta(selected_positions,
                reducer = "Johnson",
                JohnsonParam = jp,
                reducerDiscernibility = cfg$reducerDiscernibility,
                roc = FALSE,
                discrete=TRUE),
        error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
    )

    if (!is.null(ros)) {
        rec <- recalculateRules(selected_positions, ros$main, discrete=TRUE)
        sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
        results1 <- rbind(results1, data.frame(
        BRT                   = cfg$BRT,
        BRTprec               = cfg$BRTprec,
        Fraction              = cfg$Fraction,
        reducerDiscernibility = cfg$reducerDiscernibility,
        accuracyMean          = ros$quality$accuracyMean,
        nRules                = nrow(ros$main),
        nSigRules             = sig,
        stringsAsFactors = FALSE
        ))
    } else {
        results1 <- rbind(results1, data.frame(
        BRT                   = cfg$BRT,
        BRTprec               = cfg$BRTprec,
        Fraction              = cfg$Fraction,
        reducerDiscernibility = cfg$reducerDiscernibility,
        accuracyMean          = NA_real_,
        nRules                = NA_integer_,
        nSigRules             = NA_integer_,
        stringsAsFactors = FALSE
        ))
    }
}

cat("\nGrid 1 complete.\n")
```

# Cell 4: Grid 1 Results

```{r}
results1 <- results1[order(-results1$accuracyMean, na.last = TRUE), ]
rownames(results1) <- NULL

cat("=== Grid 1: Top 15 by mean accuracy ===\n\n")
print(results1[1:min(15, nrow(results1)), ], row.names = FALSE)

best1 <- results1[1, ]
cat("\nBest Grid 1 config:\n")
cat("  BRT:", best1$BRT, " BRTprec:", best1$BRTprec, " Fraction:", best1$Fraction, "\n")
cat("  Accuracy:", best1$accuracyMean, " Rules:", best1$nRules, "(", best1$nSigRules, "sig)\n")

# Store best JohnsonParam for Grid 2
best_jp <- list(
    Modulo      = TRUE,
    BRT         = best1$BRT,
    BRTprec     = best1$BRTprec,
    Precompute  = FALSE,
    Approximate = TRUE,
    Fraction    = best1$Fraction
)
best1_discern <- best1$reducerDiscernibility
```

# ============================================================
# GRID SEARCH 2: reducerDiscernibility / underSample (256 configs)
# ============================================================

# Cell 5: Build Grid 2

```{r}
classifier_vals     <- c("StandardVoter", "ObjectTrackingVoter", "NaiveBayesClassifier")
underSampleNum_vals <- c(0, 100, 250)

grid2_no_us <- expand.grid(
    classifier     = classifier_vals,
    underSample    = FALSE,
    underSampleNum = 0,
    stringsAsFactors = FALSE
)

grid2_us <- expand.grid(
    classifier     = classifier_vals,
    underSample    = TRUE,
    underSampleNum = underSampleNum_vals,
    stringsAsFactors = FALSE
)

grid2 <- rbind(grid2_no_us, grid2_us)
cat("Grid 2 configs:", nrow(grid2), "\n")
```

# Cell 6: Run Grid 2

```{r}
results2 <- data.frame(
    classifier     = character(),
    underSample    = logical(),
    underSampleNum = integer(),
    accuracyMean   = numeric(),
    nRules         = integer(),
    nSigRules      = integer(),
    stringsAsFactors = FALSE
)

for (i in 1:nrow(grid2)) {
    cfg <- grid2[i, ]
    cat(sprintf("Grid2 %d/%d — cls=%s US=%s USnum=%d\n",
                i, nrow(grid2), cfg$classifier, cfg$underSample, cfg$underSampleNum))

    ros <- tryCatch({
        if (cfg$underSample) {
        rosetta(selected_positions,
                reducer = "Johnson",
                JohnsonParam = best_jp,
                reducerDiscernibility = best1_discern,
                discrete = TRUE,
                classifier = cfg$classifier,
                underSample = TRUE,
                underSampleNum = cfg$underSampleNum,
                underSampleSize = 0,
                roc = FALSE,)
        } else {
        rosetta(selected_positions,
                reducer = "Johnson",
                JohnsonParam = best_jp,
                reducerDiscernibility = best1_discern,
                discrete = TRUE,
                classifier = cfg$classifier,
                underSample = FALSE,
                roc = FALSE)
        }
    },
    error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
    )

    rec <- recalculateRules(selected_positions, ros$main, discrete=TRUE)
    sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
    results2 <- rbind(results2, data.frame(
    classifier     = cfg$classifier,
    underSample    = cfg$underSample,
    underSampleNum = cfg$underSampleNum,
    accuracyMean   = ros$quality$accuracyMean,
    nRules         = nrow(ros$main),
    nSigRules      = sig,
    stringsAsFactors = FALSE
    ))
}

cat("\nGrid 2 complete.\n")
```

# Cell 7: Grid 2 Results

```{r}
results2 <- results2[order(-results2$accuracyMean, na.last = TRUE), ]
rownames(results2) <- NULL

cat("=== Grid 2: Top 15 by mean accuracy ===\n\n")
print(results2[1:min(15, nrow(results2)), ], row.names = FALSE)

best2 <- results2[1, ]
cat("\nBest Grid 2 config:\n")
cat("  reducerDiscernibility:", best2$reducerDiscernibility, "\n")
cat("  underSample:", best2$underSample, " underSampleNum:", best2$underSampleNum, "\n")
cat("  Accuracy:", best2$accuracyMean, " Rules:", best2$nRules, "(", best2$nSigRules, "sig)\n")

# Store best params for Grid 3
best2_classifier            <- best2$classifier
best2_underSample           <- best2$underSample
best2_underSampleNum        <- best2$underSampleNum
```

# ============================================================
# GRID SEARCH 3: Rule Filtration (126 configs)
# ============================================================

# Cell 8: Build Grid 3

```{r}
# ruleFiltration=FALSE: 1 config
# ruleFiltration=TRUE: ruleFiltrSupport × ruleFiltrAccuracy × ruleFiltrCoverage = 5×5×5 = 125
# Total: 126

# Each filter param is c(low, high). Values to try:
# Support:  c(0,0), c(1,3) [default], c(1,5), c(2,5), c(1,10)
# Accuracy: c(0,0), c(0,0.5) [default], c(0,0.7), c(0,0.85), c(0.5,1)
# Coverage: c(0,0), c(0,0) [default=same as off], c(0,0.05), c(0,0.1), c(0,0.2)

support_vals <- list(c(0,0), c(1,3), c(1,5), c(2,5), c(1,10))
accuracy_vals <- list(c(0,0), c(0,0.5), c(0,0.7), c(0,0.85), c(0.5,1))
coverage_vals <- list(c(0,0), c(0,0.05), c(0,0.1), c(0,0.2), c(0,0.5))

# Build grid as index combinations
grid3_true <- expand.grid(
    supp_idx = 1:5,
    acc_idx  = 1:5,
    cov_idx  = 1:5,
    stringsAsFactors = FALSE
)

cat("Grid 3 configs: 1 (no filtration) +", nrow(grid3_true), "(with filtration) =", 1 + nrow(grid3_true), "\n")
```

# Cell 9: Run Grid 3 — No Filtration Baseline

```{r}
run_with_best <- function(selected_positions, best_jp, best2, best1_discern, filt, filt_supp, filt_acc, filt_cov) {
  if (best2$underSample) {
    rosetta(selected_positions,
            reducer = "Johnson",
            JohnsonParam = best_jp,
            reducerDiscernibility = best1_discern,
            discrete = TRUE,
            classifier = best2$classifier,
            underSample = TRUE,
            underSampleNum = best2$underSampleNum,
            underSampleSize = 0,
            ruleFiltration = filt,
            ruleFiltrSupport = filt_supp,
            ruleFiltrAccuracy = filt_acc,
            ruleFiltrCoverage = filt_cov,
            roc = FALSE)
  } else {
    rosetta(selected_positions,
            reducer = "Johnson",
            JohnsonParam = best_jp,
            reducerDiscernibility = best1_discern,
            discrete = TRUE,
            classifier = best2$classifier,
            underSample = FALSE,
            ruleFiltration = filt,
            ruleFiltrSupport = filt_supp,
            ruleFiltrAccuracy = filt_acc,
            ruleFiltrCoverage = filt_cov,
            roc = FALSE)
  }
}

results3 <- data.frame(
  ruleFiltration   = logical(),
  filtrSupport     = character(),
  filtrAccuracy    = character(),
  filtrCoverage    = character(),
  accuracyMean     = numeric(),
  nRules           = integer(),
  nSigRules        = integer(),
  stringsAsFactors = FALSE
)

cat("Grid3 1/126 — ruleFiltration=FALSE\n")
ros <- tryCatch(
  run_with_best(selected_positions, best_jp, best2, best1_discern,
                filt = FALSE,
                filt_supp = c(0, 0),
                filt_acc  = c(0, 0),
                filt_cov  = c(0, 0)),
  error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
)

if (!is.null(ros)) {
  rec <- recalculateRules(selected_positions, ros$main, discrete=TRUE)
  sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
  results3 <- rbind(results3, data.frame(
    ruleFiltration = FALSE,
    filtrSupport   = "c(0,0)",
    filtrAccuracy  = "c(0,0)",
    filtrCoverage  = "c(0,0)",
    accuracyMean   = ros$quality$accuracyMean,
    nRules         = nrow(ros$main),
    nSigRules      = sig,
    stringsAsFactors = FALSE
  ))
} else {
  results3 <- rbind(results3, data.frame(
    ruleFiltration = FALSE,
    filtrSupport   = "c(0,0)",
    filtrAccuracy  = "c(0,0)",
    filtrCoverage  = "c(0,0)",
    accuracyMean   = NA_real_,
    nRules         = NA_integer_,
    nSigRules      = NA_integer_,
    stringsAsFactors = FALSE
  ))
}

cat("Baseline done.\n")
```

# Cell 10: Run Grid 3 — With Filtration (125 configs)

```{r}
for (i in 1:nrow(grid3_true)) {
    si <- grid3_true$supp_idx[i]
    ai <- grid3_true$acc_idx[i]
    ci <- grid3_true$cov_idx[i]

    s_val <- support_vals[[si]]
    a_val <- accuracy_vals[[ai]]
    c_val <- coverage_vals[[ci]]

    cat(sprintf("Grid3 %d/126 — Supp=c(%s) Acc=c(%s) Cov=c(%s)\n",
                i + 1,
                paste(s_val, collapse=","),
                paste(a_val, collapse=","),
                paste(c_val, collapse=",")))

    ros <- tryCatch(
        run_with_best(selected_positions, best_jp, best2, best1_discern,
                    filt = TRUE,
                    filt_supp = s_val,
                    filt_acc  = a_val,
                    filt_cov  = c_val),
        error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
    )

    if (!is.null(ros)) {
        rec <- recalculateRules(selected_positions, ros$main, discrete=TRUE)
        sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
        results3 <- rbind(results3, data.frame(
        ruleFiltration = TRUE,
        filtrSupport   = paste0("c(", paste(s_val, collapse=","), ")"),
        filtrAccuracy  = paste0("c(", paste(a_val, collapse=","), ")"),
        filtrCoverage  = paste0("c(", paste(c_val, collapse=","), ")"),
        accuracyMean   = ros$quality$accuracyMean,
        nRules         = nrow(ros$main),
        nSigRules      = sig,
        stringsAsFactors = FALSE
        ))
    } else {
        results3 <- rbind(results3, data.frame(
        ruleFiltration = TRUE,
        filtrSupport   = paste0("c(", paste(s_val, collapse=","), ")"),
        filtrAccuracy  = paste0("c(", paste(a_val, collapse=","), ")"),
        filtrCoverage  = paste0("c(", paste(c_val, collapse=","), ")"),
        accuracyMean   = NA_real_,
        nRules         = NA_integer_,
        nSigRules      = NA_integer_,
        stringsAsFactors = FALSE
        ))
    }
}

cat("\nGrid 3 complete.\n")
```

# Cell 11: Grid 3 Results

```{r}
results3 <- results3[order(-results3$accuracyMean, na.last = TRUE), ]
rownames(results3) <- NULL

cat("=== Grid 3: Top 15 by mean accuracy ===\n\n")
print(results3[1:min(15, nrow(results3)), ], row.names = FALSE)

best3 <- results3[1, ]
cat("\nBest Grid 3 config:\n")
cat("  ruleFiltration:", best3$ruleFiltration, "\n")
cat("  filtrSupport:", best3$filtrSupport, "\n")
cat("  filtrAccuracy:", best3$filtrAccuracy, "\n")
cat("  filtrCoverage:", best3$filtrCoverage, "\n")
cat("  Accuracy:", best3$accuracyMean, " Rules:", best3$nRules, "(", best3$nSigRules, "sig)\n")
```

# ============================================================
# FINAL SUMMARY
# ============================================================

# Cell 12: Print Best Overall Configuration

```{r}
cat("====================================================\n")
cat("       BEST CONFIGURATION ACROSS ALL 3 GRIDS        \n")
cat("====================================================\n\n")

cat("--- Johnson Reducer Params (Grid 1) ---\n")
cat("  Modulo:               TRUE\n")
cat("  BRT:                 ", best1$BRT, "\n")
cat("  BRTprec:             ", best1$BRTprec, "\n")
cat("  Precompute:           FALSE\n")
cat("  Approximate:          TRUE\n")
cat("  Fraction:            ", best1$Fraction, "\n")
cat("  reducerDiscernibility:", best1_discern, "\n\n")

cat("--- Discretization, Classifier & Sampling (Grid 2) ---\n")
cat("  classifier:          ", best2$classifier, "\n")
cat("  underSample:         ", best2$underSample, "\n")
if (best2$underSample) cat("  underSampleNum:      ", best2$underSampleNum, "\n")
cat("\n")

cat("--- Rule Filtration (Grid 3) ---\n")
cat("  ruleFiltration:      ", best3$ruleFiltration, "\n")
if (best3$ruleFiltration) {
    cat("  ruleFiltrSupport:  ", best3$filtrSupport, "\n")
    cat("  ruleFiltrAccuracy: ", best3$filtrAccuracy, "\n")
    cat("  ruleFiltrCoverage: ", best3$filtrCoverage, "\n")
}
cat("\n")

cat("--- Performance ---\n")
cat("  Grid 1 best accuracy:", best1$accuracyMean, "\n")
cat("  Grid 2 best accuracy:", best2$accuracyMean, "\n")
cat("  Grid 3 best accuracy:", best3$accuracyMean, "\n")
cat("====================================================\n")

best_filt       <- best3$ruleFiltration
best_filt_supp  <- eval(parse(text = best3$filtrSupport))
best_filt_acc   <- eval(parse(text = best3$filtrAccuracy))
best_filt_cov   <- eval(parse(text = best3$filtrCoverage))

best_config <- list(
    jp                    = best_jp,
    reducerDiscernibility = best1_discern,
    classifier            = best2$classifier,
    underSample           = best2$underSample,
    underSampleNum        = best2$underSampleNum,
    ruleFiltration        = best_filt,
    ruleFiltrSupport      = best_filt_supp,
    ruleFiltrAccuracy     = best_filt_acc,
    ruleFiltrCoverage     = best_filt_cov
)
```

====================================================
       BEST CONFIGURATION ACROSS ALL 3 GRIDS        
====================================================

--- Johnson Reducer Params (Grid 1) ---
  Modulo:               TRUE
  BRT:                  TRUE 
  BRTprec:              0.6 
  Precompute:           FALSE
  Approximate:          TRUE
  Fraction:             0.99 
  reducerDiscernibility: Object 

--- Discretization, Classifier & Sampling (Grid 2) ---
  classifier:           StandardVoter 
  underSample:          FALSE 

--- Rule Filtration (Grid 3) ---
  ruleFiltration:       FALSE 

--- Performance ---
  Grid 1 best accuracy: 0.962092 
  Grid 2 best accuracy: 0.962092 
  Grid 3 best accuracy: 0.962092 

# Cell 13: Genetic
```{r}
cat("=== Johnson vs Genetic on best config ===\n\n")

cat("Running Johnson...\n")
ros_johnson <- rosetta(selected_positions,
    reducer = "Johnson",
    JohnsonParam = best_config$jp,
    reducerDiscernibility = best_config$reducerDiscernibility,
    discrete = TRUE,
    classifier = best_config$classifier,
    underSample = best_config$underSample,
    underSampleNum = best_config$underSampleNum,
    underSampleSize = 0,
    ruleFiltration = best_config$ruleFiltration,
    ruleFiltrSupport = best_config$ruleFiltrSupport,
    ruleFiltrAccuracy = best_config$ruleFiltrAccuracy,
    ruleFiltrCoverage = best_config$ruleFiltrCoverage,
    roc = TRUE,
    clroc = levels(selected_positions$decision)[1])

cat("Running Genetic...\n")
gp_simple <- list(
    Modulo      = best_config$jp$Modulo,
    BRT         = best_config$jp$BRT,
    BRTprec     = best_config$jp$BRTprec,
    Precompute  = best_config$jp$Precompute,
    Approximate = best_config$jp$Approximate,
    Fraction    = best_config$jp$Fraction,
    Algorithm   = "Simple"
)

ros_genetic_simple <- rosetta(selected_positions,
    reducer = "Genetic",
    GeneticParam = gp_simple,
    reducerDiscernibility = best_config$reducerDiscernibility,
    discrete = TRUE,
    classifier = best_config$classifier,
    underSample = best_config$underSample,
    underSampleNum = best_config$underSampleNum,
    underSampleSize = 0,
    ruleFiltration = best_config$ruleFiltration,
    ruleFiltrSupport = best_config$ruleFiltrSupport,
    ruleFiltrAccuracy = best_config$ruleFiltrAccuracy,
    ruleFiltrCoverage = best_config$ruleFiltrCoverage,
    roc = TRUE,
    clroc = levels(selected_positions$decision)[1])

cat("\n--- Johnson ---\n")
print(ros_johnson$quality)
cat("Rules:", nrow(ros_johnson$main), "\n")

cat("\n--- Genetic ---\n")
print(ros_genetic_simple$quality)
cat("Rules:", nrow(ros_genetic_simple$main), "\n")
```