---
title: "3.3 - Parameters Grid Search"
output: html_document
---

# Cell 1: Setup and Data

```{r}
set.seed(435)
library(R.ROSETTA)

selected_positions <- read.csv("data/clean/selected_positions.csv")

cat("Data:", nrow(selected_positions), "rows x", ncol(selected_positions), "cols\n")
cat("Classes:", paste(names(table(selected_positions$class)), table(selected_positions$class), sep="=", collapse=", "), "\n")
```

# ============================================================
# GRID SEARCH 1: Johnson BRT / BRTprec / Fraction (64 configs)
# ============================================================

# Cell 2: Build Grid 1

```{r}
fraction_vals <- c(0.99, 0.95, 0.9, 0.8, 0.7, 0.5, 0.25, 0.1)
brtprec_vals  <- c(1, 0.9, 0.8, 0.6, 0.5, 0.25, 0.1)

grid1_brt_false <- expand.grid(
    BRT     = FALSE,
    BRTprec = 0,
    Fraction = fraction_vals,
    stringsAsFactors = FALSE
)

grid1_brt_true <- expand.grid(
    BRT     = TRUE,
    BRTprec = brtprec_vals,
    Fraction = fraction_vals,
    stringsAsFactors = FALSE
)

grid1 <- rbind(grid1_brt_false, grid1_brt_true)
cat("Grid 1 configs:", nrow(grid1), "\n")
```

# Cell 3: Run Grid 1

```{r}
results1 <- data.frame(
    BRT          = logical(),
    BRTprec      = numeric(),
    Fraction     = numeric(),
    accuracyMean = numeric(),
    nRules       = integer(),
    nSigRules    = integer(),
    stringsAsFactors = FALSE
)

i <- 3
cfg <- grid1[i, ]
cat(sprintf("Grid1 %d/%d — BRT=%s BRTprec=%.2f Fraction=%.2f\n",
            i, nrow(grid1), cfg$BRT, cfg$BRTprec, cfg$Fraction))

for (i in 1:nrow(grid1)) {
    cfg <- grid1[i, ]
    cat(sprintf("Grid1 %d/%d — BRT=%s BRTprec=%.2f Fraction=%.2f\n",
                i, nrow(grid1), cfg$BRT, cfg$BRTprec, cfg$Fraction))

    jp <- list(
        Modulo      = TRUE,
        BRT         = cfg$BRT,
        BRTprec     = cfg$BRTprec,
        Precompute  = FALSE,
        Approximate = TRUE,
        Fraction    = cfg$Fraction
    )

    ros <- tryCatch(
        rosetta(selected_positions,
                reducer = "Johnson",
                JohnsonParam = jp,
                roc = FALSE,
                discrete = TRUE),
        error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
    )
    rec <- recalculateRules(selected_positions, ros$main, discrete=TRUE)

    sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
    results1 <- rbind(results1, data.frame(
        BRT          = cfg$BRT,
        BRTprec      = cfg$BRTprec,
        Fraction     = cfg$Fraction,
        accuracyMean = ros$quality$accuracyMean,
        nRules       = nrow(ros$main),
        nSigRules    = sig,
        stringsAsFactors = FALSE
    ))
}

cat("\nGrid 1 complete.\n")
```

# Cell 4: Grid 1 Results

```{r}
results1 <- results1[order(-results1$accuracyMean, na.last = TRUE), ]
rownames(results1) <- NULL

cat("=== Grid 1: Top 15 by mean accuracy ===\n\n")
print(results1[1:min(15, nrow(results1)), ], row.names = FALSE)

best1 <- results1[1, ]
cat("\nBest Grid 1 config:\n")
cat("  BRT:", best1$BRT, " BRTprec:", best1$BRTprec, " Fraction:", best1$Fraction, "\n")
cat("  Accuracy:", best1$accuracyMean, " Rules:", best1$nRules, "(", best1$nSigRules, "sig)\n")

best_jp <- list(
    Modulo      = TRUE,
    BRT         = best1$BRT,
    BRTprec     = best1$BRTprec,
    Precompute  = FALSE,
    Approximate = TRUE,
    Fraction    = best1$Fraction
)
```

# ============================================================
# GRID SEARCH 2: reducerDiscernibility / underSample (256 configs)
# ============================================================

# Cell 5: Build Grid 2

```{r}
classifier_vals <- c("StandardVoter", "ObjectTrackingVoter", "NaiveBayesClassifier")
discern_vals    <- c("Full", "Object")

grid2 <- expand.grid(
    classifier            = classifier_vals,
    reducerDiscernibility = discern_vals,
    stringsAsFactors = FALSE
)

cat("Grid 2 configs:", nrow(grid2), "\n")
```

# Cell 6: Run Grid 2

```{r}
results2 <- data.frame(
    classifier            = character(),
    reducerDiscernibility = character(),
    accuracyMean          = numeric(),
    nRules                = integer(),
    nSigRules             = integer(),
    stringsAsFactors = FALSE
)

for (i in 1:nrow(grid2)) {
    cfg <- grid2[i, ]
    cat(sprintf("Grid2 %d/%d — cls=%s discern=%s\n",
                i, nrow(grid2), cfg$classifier, cfg$reducerDiscernibility))

    ros <- tryCatch(
        suppressWarnings(rosetta(selected_positions,
                reducer = "Johnson",
                JohnsonParam = best_jp,
                reducerDiscernibility = cfg$reducerDiscernibility,
                discrete = TRUE,
                classifier = cfg$classifier,
                roc = FALSE)),
        error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
    )
    rec <- recalculateRules(selected_positions, ros$main, discrete=TRUE)

    sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
    results2 <- rbind(results2, data.frame(
        classifier            = cfg$classifier,
        reducerDiscernibility = cfg$reducerDiscernibility,
        accuracyMean          = ros$quality$accuracyMean,
        nRules                = nrow(ros$main),
        nSigRules             = sig,
        stringsAsFactors = FALSE
    ))
}

cat("\nGrid 2 complete.\n")
```

# Cell 7: Grid 2 Results

```{r}
results2 <- results2[order(-results2$accuracyMean, na.last = TRUE), ]
rownames(results2) <- NULL

cat("=== Grid 2: Top 15 by mean accuracy ===\n\n")
print(results2[1:min(15, nrow(results2)), ], row.names = FALSE)

best2 <- results2[1, ]
cat("\nBest Grid 2 config:\n")
cat("  classifier:", best2$classifier, "\n")
cat("  reducerDiscernibility:", best2$reducerDiscernibility, "\n")
cat("  Accuracy:", best2$accuracyMean, " Rules:", best2$nRules, "(", best2$nSigRules, "sig)\n")

best2_classifier <- best2$classifier
best2_discern    <- best2$reducerDiscernibility
```

# ============================================================
# GRID SEARCH 3: Rule Filtration (126 configs)
# ============================================================

# Cell 8: Build Grid 3

```{r}
support_vals <- list(c(0,0), c(1,3), c(1,5), c(2,5), c(1,10))
accuracy_vals <- list(c(0,0), c(0,0.5), c(0,0.7), c(0,0.85), c(0.5,1))
coverage_vals <- list(c(0,0), c(0,0.05), c(0,0.1), c(0,0.2), c(0,0.5))

# No filtration: 1 config
grid3_no_filt <- data.frame(
    ruleFiltration = FALSE,
    supp_idx = 1,
    acc_idx = 1,
    cov_idx = 1,
    stringsAsFactors = FALSE
)

# With filtration: 5×5×5 = 125 configs
grid3_with_filt <- expand.grid(
    supp_idx = 1:5,
    acc_idx  = 1:5,
    cov_idx  = 1:5,
    stringsAsFactors = FALSE
)
grid3_with_filt$ruleFiltration <- TRUE

grid3 <- rbind(grid3_no_filt, grid3_with_filt)
cat("Grid 3 configs:", nrow(grid3), "\n")
```

# Cell 9: Run Grid 3

```{r}
results3 <- data.frame(
    ruleFiltration   = logical(),
    filtrSupport     = character(),
    filtrAccuracy    = character(),
    filtrCoverage    = character(),
    accuracyMean     = numeric(),
    nRules           = integer(),
    nSigRules        = integer(),
    stringsAsFactors = FALSE
)

for (i in 1:nrow(grid3)) {
    cfg <- grid3[i, ]
    
    s_val <- support_vals[[cfg$supp_idx]]
    a_val <- accuracy_vals[[cfg$acc_idx]]
    c_val <- coverage_vals[[cfg$cov_idx]]
    
    cat(sprintf("Grid3 %d/%d — Filt=%s Supp=c(%s) Acc=c(%s) Cov=c(%s)\n",
                i, nrow(grid3),
                cfg$ruleFiltration,
                paste(s_val, collapse=","),
                paste(a_val, collapse=","),
                paste(c_val, collapse=",")))
    
    ros <- tryCatch(
        rosetta(selected_positions,
                reducer = "Johnson",
                JohnsonParam = best_jp,
                reducerDiscernibility = best2_discern,
                discrete = TRUE,
                classifier = best2_classifier,
                ruleFiltration = cfg$ruleFiltration,
                ruleFiltrSupport = s_val,
                ruleFiltrAccuracy = a_val,
                ruleFiltrCoverage = c_val,
                roc = FALSE),
        error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
    )
    rec <- recalculateRules(selected_positions, ros$main, discrete=TRUE)
    sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
    results3 <- rbind(results3, data.frame(
        ruleFiltration = cfg$ruleFiltration,
        filtrSupport   = paste0("c(", paste(s_val, collapse=","), ")"),
        filtrAccuracy  = paste0("c(", paste(a_val, collapse=","), ")"),
        filtrCoverage  = paste0("c(", paste(c_val, collapse=","), ")"),
        accuracyMean   = ros$quality$accuracyMean,
        nRules         = nrow(ros$main),
        nSigRules      = sig,
        stringsAsFactors = FALSE
    ))
}

cat("\nGrid 3 complete.\n")
```

# Cell 10: Grid 3 Results

```{r}
results3 <- results3[order(-results3$accuracyMean, na.last = TRUE), ]
rownames(results3) <- NULL

cat("=== Grid 3: Top 15 by mean accuracy ===\n\n")
print(results3[1:min(15, nrow(results3)), ], row.names = FALSE)

best3 <- results3[1, ]
cat("\nBest Grid 3 config:\n")
cat("  ruleFiltration:", best3$ruleFiltration, "\n")
cat("  filtrSupport:", best3$filtrSupport, "\n")
cat("  filtrAccuracy:", best3$filtrAccuracy, "\n")
cat("  filtrCoverage:", best3$filtrCoverage, "\n")
cat("  Accuracy:", best3$accuracyMean, " Rules:", best3$nRules, "(", best3$nSigRules, "sig)\n")

best_filt       <- best3$ruleFiltration
best_filt_supp  <- eval(parse(text = best3$filtrSupport))
best_filt_acc   <- eval(parse(text = best3$filtrAccuracy))
best_filt_cov   <- eval(parse(text = best3$filtrCoverage))
```

# ============================================================
# FINAL SUMMARY
# ============================================================

# Cell 11: Print Best Overall Configuration

```{r}
cat("====================================================\n")
cat("       BEST CONFIGURATION ACROSS ALL 3 GRIDS        \n")
cat("====================================================\n\n")

cat("--- Johnson Reducer Params (Grid 1) ---\n")
cat("  Modulo:               TRUE\n")
cat("  BRT:                 ", best1$BRT, "\n")
cat("  BRTprec:             ", best1$BRTprec, "\n")
cat("  Precompute:           FALSE\n")
cat("  Approximate:          TRUE\n")
cat("  Fraction:            ", best1$Fraction, "\n\n")

cat("--- Discretization, Classifier & Sampling (Grid 2) ---\n")
cat("  classifier:          ", best2$classifier, "\n")
cat("  reducerDiscernibility:", best2_discern, "\n")
cat("\n")

cat("--- Rule Filtration (Grid 3) ---\n")
cat("  ruleFiltration:      ", best3$ruleFiltration, "\n")
if (best3$ruleFiltration) {
    cat("  ruleFiltrSupport:  ", best3$filtrSupport, "\n")
    cat("  ruleFiltrAccuracy: ", best3$filtrAccuracy, "\n")
    cat("  ruleFiltrCoverage: ", best3$filtrCoverage, "\n")
}
cat("\n")

cat("--- Performance ---\n")
cat("  Grid 1 best accuracy:", best1$accuracyMean, "\n")
cat("  Grid 2 best accuracy:", best2$accuracyMean, "\n")
cat("  Grid 3 best accuracy:", best3$accuracyMean, "\n")
cat("====================================================\n")

best_filt       <- best3$ruleFiltration
best_filt_supp  <- eval(parse(text = best3$filtrSupport))
best_filt_acc   <- eval(parse(text = best3$filtrAccuracy))
best_filt_cov   <- eval(parse(text = best3$filtrCoverage))

best_config <- list(
    jp                    = best_jp,
    reducerDiscernibility = best2_discern,
    classifier            = best2_classifier,
    ruleFiltration        = best_filt,
    ruleFiltrSupport      = best_filt_supp,
    ruleFiltrAccuracy     = best_filt_acc,
    ruleFiltrCoverage     = best_filt_cov
)
```

====================================================
       BEST CONFIGURATION ACROSS ALL 3 GRIDS        
====================================================

--- Johnson Reducer Params (Grid 1) ---
  Modulo:               TRUE
  BRT:                  TRUE 
  BRTprec:              0.9 
  Precompute:           FALSE
  Approximate:          TRUE
  Fraction:             0.99 

--- Discretization, Classifier & Sampling (Grid 2) ---
  classifier:           StandardVoter 
  reducerDiscernibility: Object 

--- Rule Filtration (Grid 3) ---
  ruleFiltration:       FALSE 

--- Performance ---
  Grid 1 best accuracy: 0.962092 
  Grid 2 best accuracy: 0.962092 
  Grid 3 best accuracy: 0.962092 

# Cell 12: Genetic
```{r}
cat("=== Johnson vs Genetic on best config ===\n\n")

cat("Running Johnson...\n")
ros_johnson <- rosetta(selected_positions,
    reducer = "Johnson",
    JohnsonParam = best_config$jp,
    reducerDiscernibility = best_config$reducerDiscernibility,
    discrete = TRUE,
    classifier = best_config$classifier,
    ruleFiltration = best_config$ruleFiltration,
    ruleFiltrSupport = best_config$ruleFiltrSupport,
    ruleFiltrAccuracy = best_config$ruleFiltrAccuracy,
    ruleFiltrCoverage = best_config$ruleFiltrCoverage,
    roc = TRUE,
    clroc = levels(selected_positions$decision)[1])

cat("Running Genetic...\n")
gp_simple <- list(
    Modulo      = best_config$jp$Modulo,
    BRT         = best_config$jp$BRT,
    BRTprec     = best_config$jp$BRTprec,
    Precompute  = best_config$jp$Precompute,
    Approximate = best_config$jp$Approximate,
    Fraction    = best_config$jp$Fraction,
    Algorithm   = "Simple"
)

ros_genetic_simple <- rosetta(selected_positions,
    reducer = "Genetic",
    GeneticParam = gp_simple,
    reducerDiscernibility = best_config$reducerDiscernibility,
    discrete = TRUE,
    classifier = best_config$classifier,
    ruleFiltration = best_config$ruleFiltration,
    ruleFiltrSupport = best_config$ruleFiltrSupport,
    ruleFiltrAccuracy = best_config$ruleFiltrAccuracy,
    ruleFiltrCoverage = best_config$ruleFiltrCoverage,
    roc = TRUE,
    clroc = levels(selected_positions$decision)[1])

cat("\n--- Johnson ---\n")
print(ros_johnson$quality)
cat("Rules:", nrow(ros_johnson$main), "\n")

cat("\n--- Genetic ---\n")
print(ros_genetic_simple$quality)
cat("Rules:", nrow(ros_genetic_simple$main), "\n")
```