---
title: "Parameters Grid Search"
output: html_document
---

# Cell 1: Setup and Data

```{r}
set.seed(435)
library(R.ROSETTA)

selected_positions <- read.csv("data/clean/selected_positions.csv")

cat("Data:", nrow(selected_positions), "rows x", ncol(selected_positions), "cols\n")
cat("Classes:", paste(names(table(selected_positions$class)), table(selected_positions$class), sep="=", collapse=", "), "\n")
```

# ============================================================
# GRID SEARCH 1: Johnson BRT / BRTprec / Fraction (64 configs)
# ============================================================

# Build Grid 1

```{r}
# BRT=FALSE: BRTprec irrelevant → use 0 as placeholder, 1 row × 8 Fraction = 8
# BRT=TRUE: 7 BRTprec values × 8 Fraction = 56
# Total: 64

fraction_vals <- c(0.99, 0.95, 0.9, 0.8, 0.7, 0.5, 0.25, 0.1)
brtprec_vals  <- c(1, 0.9, 0.8, 0.6, 0.5, 0.25, 0.1)

grid1_brt_false <- expand.grid(
  BRT     = FALSE,
  BRTprec = 0,
  Fraction = fraction_vals,
  stringsAsFactors = FALSE
)

grid1_brt_true <- expand.grid(
  BRT     = TRUE,
  BRTprec = brtprec_vals,
  Fraction = fraction_vals,
  stringsAsFactors = FALSE
)

grid1 <- rbind(grid1_brt_false, grid1_brt_true)
cat("Grid 1 configs:", nrow(grid1), "\n")
```

# Run Grid 1

```{r}
results1 <- data.frame(
  BRT          = logical(),
  BRTprec      = numeric(),
  Fraction     = numeric(),
  accuracyMean = numeric(),
  nRules       = integer(),
  nSigRules    = integer(),
  stringsAsFactors = FALSE
)

for (i in 1:nrow(grid1)) {
  cfg <- grid1[i, ]
  cat(sprintf("Grid1 %d/%d — BRT=%s BRTprec=%.2f Fraction=%.2f\n",
              i, nrow(grid1), cfg$BRT, cfg$BRTprec, cfg$Fraction))

  jp <- list(
    Modulo      = TRUE,
    BRT         = cfg$BRT,
    BRTprec     = cfg$BRTprec,
    Precompute  = FALSE,
    Approximate = TRUE,
    Fraction    = cfg$Fraction
  )

  ros <- tryCatch(
    rosetta(selected_positions,
            reducer = "Johnson",
            JohnsonParam = jp,
            roc = FALSE,
            seed = 435),
    error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
  )

  if (!is.null(ros)) {
    rec <- recalculateRules(selected_positions, ros$main)
    sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
    results1 <- rbind(results1, data.frame(
      BRT          = cfg$BRT,
      BRTprec      = cfg$BRTprec,
      Fraction     = cfg$Fraction,
      accuracyMean = ros$quality$accuracyMean,
      nRules       = nrow(ros$main),
      nSigRules    = sig,
      stringsAsFactors = FALSE
    ))
  } else {
    results1 <- rbind(results1, data.frame(
      BRT          = cfg$BRT,
      BRTprec      = cfg$BRTprec,
      Fraction     = cfg$Fraction,
      accuracyMean = NA_real_,
      nRules       = NA_integer_,
      nSigRules    = NA_integer_,
      stringsAsFactors = FALSE
    ))
  }
}

cat("\nGrid 1 complete.\n")
```

# Grid 1 Results

```{r}
results1 <- results1[order(-results1$accuracyMean, na.last = TRUE), ]
rownames(results1) <- NULL

cat("=== Grid 1: Top 5 by mean accuracy ===\n\n")
print(results1[1:min(5, nrow(results1)), ], row.names = FALSE)

best1 <- results1[1, ]
cat("\nBest Grid 1 config:\n")
cat("  BRT:", best1$BRT, " BRTprec:", best1$BRTprec, " Fraction:", best1$Fraction, "\n")
cat("  Accuracy:", best1$accuracyMean, " Rules:", best1$nRules, "(", best1$nSigRules, "sig)\n")

# Store best JohnsonParam for Grid 2
best_jp <- list(
  Modulo      = TRUE,
  BRT         = best1$BRT,
  BRTprec     = best1$BRTprec,
  Precompute  = FALSE,
  Approximate = TRUE,
  Fraction    = best1$Fraction
)
```

# ============================================================
# GRID SEARCH 2: DiscreteMethod / DiscreteParam / reducerDiscernibility / underSample (256 configs)
# ============================================================

# Build Grid 2

```{r}
# underSample=FALSE: underSampleNum irrelevant → 1 value, 4×8×2×1 = 64
# underSample=TRUE: underSampleNum has 3 values, 4×8×2×3 = 192
# Total: 256

discreteMethod_vals       <- c("EqualFrequency", "MDL", "Naive", "SemiNaive")
discreteParam_vals        <- c(2, 3, 4, 5, 6, 7, 8, 10)
reducerDiscernibility_vals <- c("Full", "Object")
underSampleNum_vals       <- c(0, 100, 250)

grid2_no_us <- expand.grid(
  discreteMethod        = discreteMethod_vals,
  discreteParam         = discreteParam_vals,
  reducerDiscernibility = reducerDiscernibility_vals,
  underSample           = FALSE,
  underSampleNum        = 0,
  stringsAsFactors = FALSE
)

grid2_us <- expand.grid(
  discreteMethod        = discreteMethod_vals,
  discreteParam         = discreteParam_vals,
  reducerDiscernibility = reducerDiscernibility_vals,
  underSample           = TRUE,
  underSampleNum        = underSampleNum_vals,
  stringsAsFactors = FALSE
)

grid2 <- rbind(grid2_no_us, grid2_us)
cat("Grid 2 configs:", nrow(grid2), "\n")
```

# Run Grid 2

```{r}
results2 <- data.frame(
  discreteMethod        = character(),
  discreteParam         = integer(),
  reducerDiscernibility = character(),
  underSample           = logical(),
  underSampleNum        = integer(),
  accuracyMean          = numeric(),
  nRules                = integer(),
  nSigRules             = integer(),
  stringsAsFactors = FALSE
)

for (i in 1:nrow(grid2)) {
  cfg <- grid2[i, ]
  cat(sprintf("Grid2 %d/%d — %s param=%d disc=%s US=%s USnum=%d\n",
              i, nrow(grid2), cfg$discreteMethod, cfg$discreteParam,
              cfg$reducerDiscernibility, cfg$underSample, cfg$underSampleNum))

  ros <- tryCatch({
    if (cfg$underSample) {
      rosetta(selected_positions,
              reducer = "Johnson",
              JohnsonParam = best_jp,
              discreteMethod = cfg$discreteMethod,
              discreteParam = cfg$discreteParam,
              reducerDiscernibility = cfg$reducerDiscernibility,
              underSample = TRUE,
              underSampleNum = cfg$underSampleNum,
              underSampleSize = 0,
              roc = FALSE,
              seed = 435)
    } else {
      rosetta(selected_positions,
              reducer = "Johnson",
              JohnsonParam = best_jp,
              discreteMethod = cfg$discreteMethod,
              discreteParam = cfg$discreteParam,
              reducerDiscernibility = cfg$reducerDiscernibility,
              underSample = FALSE,
              roc = FALSE,
              seed = 435)
    }
  },
  error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
  )

  if (!is.null(ros)) {
    rec <- recalculateRules(selected_positions, ros$main)
    sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
    results2 <- rbind(results2, data.frame(
      discreteMethod        = cfg$discreteMethod,
      discreteParam         = cfg$discreteParam,
      reducerDiscernibility = cfg$reducerDiscernibility,
      underSample           = cfg$underSample,
      underSampleNum        = cfg$underSampleNum,
      accuracyMean          = ros$quality$accuracyMean,
      nRules                = nrow(ros$main),
      nSigRules             = sig,
      stringsAsFactors = FALSE
    ))
  } else {
    results2 <- rbind(results2, data.frame(
      discreteMethod        = cfg$discreteMethod,
      discreteParam         = cfg$discreteParam,
      reducerDiscernibility = cfg$reducerDiscernibility,
      underSample           = cfg$underSample,
      underSampleNum        = cfg$underSampleNum,
      accuracyMean          = NA_real_,
      nRules                = NA_integer_,
      nSigRules             = NA_integer_,
      stringsAsFactors = FALSE
    ))
  }
}

cat("\nGrid 2 complete.\n")
```

# Grid 2 Results

```{r}
results2 <- results2[order(-results2$accuracyMean, na.last = TRUE), ]
rownames(results2) <- NULL

cat("=== Grid 2: Top 5 by mean accuracy ===\n\n")
print(results2[1:min(5, nrow(results2)), ], row.names = FALSE)

best2 <- results2[1, ]
cat("\nBest Grid 2 config:\n")
cat("  discreteMethod:", best2$discreteMethod, "\n")
cat("  discreteParam:", best2$discreteParam, "\n")
cat("  reducerDiscernibility:", best2$reducerDiscernibility, "\n")
cat("  underSample:", best2$underSample, " underSampleNum:", best2$underSampleNum, "\n")
cat("  Accuracy:", best2$accuracyMean, " Rules:", best2$nRules, "(", best2$nSigRules, "sig)\n")

# Store best params for Grid 3
best2_discreteMethod        <- best2$discreteMethod
best2_discreteParam         <- best2$discreteParam
best2_reducerDiscernibility <- best2$reducerDiscernibility
best2_underSample           <- best2$underSample
best2_underSampleNum        <- best2$underSampleNum
```

# ============================================================
# GRID SEARCH 3: Rule Filtration (126 configs)
# ============================================================

# Build Grid 3

```{r}
# ruleFiltration=FALSE: 1 config
# ruleFiltration=TRUE: ruleFiltrSupport × ruleFiltrAccuracy × ruleFiltrCoverage = 5×5×5 = 125
# Total: 126

# Each filter param is c(low, high). Values to try:
# Support:  c(0,0), c(1,3) [default], c(1,5), c(2,5), c(1,10)
# Accuracy: c(0,0), c(0,0.5) [default], c(0,0.7), c(0,0.85), c(0.5,1)
# Coverage: c(0,0), c(0,0) [default=same as off], c(0,0.05), c(0,0.1), c(0,0.2)

support_vals <- list(c(0,0), c(1,3), c(1,5), c(2,5), c(1,10))
accuracy_vals <- list(c(0,0), c(0,0.5), c(0,0.7), c(0,0.85), c(0.5,1))
coverage_vals <- list(c(0,0), c(0,0.05), c(0,0.1), c(0,0.2), c(0,0.5))

# Build grid as index combinations
grid3_true <- expand.grid(
  supp_idx = 1:5,
  acc_idx  = 1:5,
  cov_idx  = 1:5,
  stringsAsFactors = FALSE
)

cat("Grid 3 configs: 1 (no filtration) +", nrow(grid3_true), "(with filtration) =", 1 + nrow(grid3_true), "\n")
```

# Run Grid 3 — No Filtration Baseline

```{r}
# Helper to run rosetta with best params from Grid 1 + Grid 2
run_with_best <- function(selected_positions, best_jp, best2, filt, filt_supp, filt_acc, filt_cov) {
  if (best2$underSample) {
    rosetta(selected_positions,
            reducer = "Johnson",
            JohnsonParam = best_jp,
            discreteMethod = best2$discreteMethod,
            discreteParam = best2$discreteParam,
            reducerDiscernibility = best2$reducerDiscernibility,
            underSample = TRUE,
            underSampleNum = best2$underSampleNum,
            underSampleSize = 0,
            ruleFiltration = filt,
            ruleFiltrSupport = filt_supp,
            ruleFiltrAccuracy = filt_acc,
            ruleFiltrCoverage = filt_cov,
            roc = FALSE,
            seed = 435)
  } else {
    rosetta(selected_positions,
            reducer = "Johnson",
            JohnsonParam = best_jp,
            discreteMethod = best2$discreteMethod,
            discreteParam = best2$discreteParam,
            reducerDiscernibility = best2$reducerDiscernibility,
            underSample = FALSE,
            ruleFiltration = filt,
            ruleFiltrSupport = filt_supp,
            ruleFiltrAccuracy = filt_acc,
            ruleFiltrCoverage = filt_cov,
            roc = FALSE,
            seed = 435)
  }
}

results3 <- data.frame(
  ruleFiltration   = logical(),
  filtrSupport     = character(),
  filtrAccuracy    = character(),
  filtrCoverage    = character(),
  accuracyMean     = numeric(),
  nRules           = integer(),
  nSigRules        = integer(),
  stringsAsFactors = FALSE
)

# Run baseline: no filtration
cat("Grid3 1/126 — ruleFiltration=FALSE\n")
ros <- tryCatch(
  run_with_best(selected_positions, best_jp, best2,
                filt = FALSE,
                filt_supp = c(0, 0),
                filt_acc  = c(0, 0),
                filt_cov  = c(0, 0)),
  error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
)

if (!is.null(ros)) {
  rec <- recalculateRules(selected_positions, ros$main)
  sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
  results3 <- rbind(results3, data.frame(
    ruleFiltration = FALSE,
    filtrSupport   = "c(0,0)",
    filtrAccuracy  = "c(0,0)",
    filtrCoverage  = "c(0,0)",
    accuracyMean   = ros$quality$accuracyMean,
    nRules         = nrow(ros$main),
    nSigRules      = sig,
    stringsAsFactors = FALSE
  ))
} else {
  results3 <- rbind(results3, data.frame(
    ruleFiltration = FALSE,
    filtrSupport   = "c(0,0)",
    filtrAccuracy  = "c(0,0)",
    filtrCoverage  = "c(0,0)",
    accuracyMean   = NA_real_,
    nRules         = NA_integer_,
    nSigRules      = NA_integer_,
    stringsAsFactors = FALSE
  ))
}

cat("Baseline done.\n")
```

# Run Grid 3 — With Filtration (125 configs)

```{r}
for (i in 1:nrow(grid3_true)) {
  si <- grid3_true$supp_idx[i]
  ai <- grid3_true$acc_idx[i]
  ci <- grid3_true$cov_idx[i]

  s_val <- support_vals[[si]]
  a_val <- accuracy_vals[[ai]]
  c_val <- coverage_vals[[ci]]

  cat(sprintf("Grid3 %d/126 — Supp=c(%s) Acc=c(%s) Cov=c(%s)\n",
              i + 1,
              paste(s_val, collapse=","),
              paste(a_val, collapse=","),
              paste(c_val, collapse=",")))

  ros <- tryCatch(
    run_with_best(selected_positions, best_jp, best2,
                  filt = TRUE,
                  filt_supp = s_val,
                  filt_acc  = a_val,
                  filt_cov  = c_val),
    error = function(e) { cat("  ERROR:", e$message, "\n"); NULL }
  )

  if (!is.null(ros)) {
    rec <- recalculateRules(selected_positions, ros$main)
    sig <- sum(rec$pValue < 0.05, na.rm = TRUE)
    results3 <- rbind(results3, data.frame(
      ruleFiltration = TRUE,
      filtrSupport   = paste0("c(", paste(s_val, collapse=","), ")"),
      filtrAccuracy  = paste0("c(", paste(a_val, collapse=","), ")"),
      filtrCoverage  = paste0("c(", paste(c_val, collapse=","), ")"),
      accuracyMean   = ros$quality$accuracyMean,
      nRules         = nrow(ros$main),
      nSigRules      = sig,
      stringsAsFactors = FALSE
    ))
  } else {
    results3 <- rbind(results3, data.frame(
      ruleFiltration = TRUE,
      filtrSupport   = paste0("c(", paste(s_val, collapse=","), ")"),
      filtrAccuracy  = paste0("c(", paste(a_val, collapse=","), ")"),
      filtrCoverage  = paste0("c(", paste(c_val, collapse=","), ")"),
      accuracyMean   = NA_real_,
      nRules         = NA_integer_,
      nSigRules      = NA_integer_,
      stringsAsFactors = FALSE
    ))
  }
}

cat("\nGrid 3 complete.\n")
```

# Grid 3 Results

```{r}
results3 <- results3[order(-results3$accuracyMean, na.last = TRUE), ]
rownames(results3) <- NULL

cat("=== Grid 3: Top 5 by mean accuracy ===\n\n")
print(results3[1:min(5, nrow(results3)), ], row.names = FALSE)

best3 <- results3[1, ]
cat("\nBest Grid 3 config:\n")
cat("  ruleFiltration:", best3$ruleFiltration, "\n")
cat("  filtrSupport:", best3$filtrSupport, "\n")
cat("  filtrAccuracy:", best3$filtrAccuracy, "\n")
cat("  filtrCoverage:", best3$filtrCoverage, "\n")
cat("  Accuracy:", best3$accuracyMean, " Rules:", best3$nRules, "(", best3$nSigRules, "sig)\n")
```

# ============================================================
# FINAL SUMMARY
# ============================================================

# Print Best Overall Configuration

```{r}
cat("====================================================\n")
cat("       BEST CONFIGURATION ACROSS ALL 3 GRIDS        \n")
cat("====================================================\n\n")

cat("--- Johnson Reducer Params (Grid 1) ---\n")
cat("  Modulo:      TRUE\n")
cat("  BRT:        ", best1$BRT, "\n")
cat("  BRTprec:    ", best1$BRTprec, "\n")
cat("  Precompute:  FALSE\n")
cat("  Approximate: TRUE\n")
cat("  Fraction:   ", best1$Fraction, "\n\n")

cat("--- Discretization & Structure (Grid 2) ---\n")
cat("  discreteMethod:       ", best2$discreteMethod, "\n")
cat("  discreteParam:        ", best2$discreteParam, "\n")
cat("  reducerDiscernibility:", best2$reducerDiscernibility, "\n")
cat("  underSample:          ", best2$underSample, "\n")
if (best2$underSample) cat("  underSampleNum:       ", best2$underSampleNum, "\n")
cat("\n")

cat("--- Rule Filtration (Grid 3) ---\n")
cat("  ruleFiltration: ", best3$ruleFiltration, "\n")
if (best3$ruleFiltration) {
  cat("  ruleFiltrSupport: ", best3$filtrSupport, "\n")
  cat("  ruleFiltrAccuracy:", best3$filtrAccuracy, "\n")
  cat("  ruleFiltrCoverage:", best3$filtrCoverage, "\n")
}
cat("\n")

cat("--- Performance ---\n")
cat("  Grid 1 best accuracy:", best1$accuracyMean, "\n")
cat("  Grid 2 best accuracy:", best2$accuracyMean, "\n")
cat("  Grid 3 best accuracy:", best3$accuracyMean, "\n")
cat("====================================================\n")

# Parse best filtration params back to numeric vectors
best_filt       <- best3$ruleFiltration
best_filt_supp  <- eval(parse(text = best3$filtrSupport))
best_filt_acc   <- eval(parse(text = best3$filtrAccuracy))
best_filt_cov   <- eval(parse(text = best3$filtrCoverage))

best_config <- list(
  jp = best_jp,
  discreteMethod        = best2$discreteMethod,
  discreteParam         = best2$discreteParam,
  reducerDiscernibility = best2$reducerDiscernibility,
  underSample           = best2$underSample,
  underSampleNum        = best2$underSampleNum,
  ruleFiltration        = best_filt,
  ruleFiltrSupport      = best_filt_supp,
  ruleFiltrAccuracy     = best_filt_acc,
  ruleFiltrCoverage     = best_filt_cov
)
```

# Cell 13: Genetic

```{r}
cat("=== Johnson vs Genetic on best config ===\n\n")

# --- Johnson ---
cat("Running Johnson...\n")
ros_johnson <- rosetta(selected_positions,
    reducer = "Johnson",
    JohnsonParam = best_config$jp,
    discreteMethod = best_config$discreteMethod,
    discreteParam = best_config$discreteParam,
    reducerDiscernibility = best_config$reducerDiscernibility,
    underSample = best_config$underSample,
    underSampleNum = best_config$underSampleNum,
    underSampleSize = 0,
    ruleFiltration = best_config$ruleFiltration,
    ruleFiltrSupport = best_config$ruleFiltrSupport,
    ruleFiltrAccuracy = best_config$ruleFiltrAccuracy,
    ruleFiltrCoverage = best_config$ruleFiltrCoverage,
    roc = TRUE,
    clroc = levels(selected_positions$class)[1],
    seed = 435)

# --- Genetic (same params, swap reducer + GeneticParam) ---
cat("Running Genetic...\n")
gp <- list(
  Modulo      = best_config$jp$Modulo,
  BRT         = best_config$jp$BRT,
  BRTprec     = best_config$jp$BRTprec,
  Precompute  = best_config$jp$Precompute,
  Approximate = best_config$jp$Approximate,
  Fraction    = best_config$jp$Fraction,
  Algorithm   = "Simple"
)

ros_genetic <- rosetta(selected_positions,
    reducer = "Genetic",
    GeneticParam = gp,
    discreteMethod = best_config$discreteMethod,
    discreteParam = best_config$discreteParam,
    reducerDiscernibility = best_config$reducerDiscernibility,
    underSample = best_config$underSample,
    underSampleNum = best_config$underSampleNum,
    underSampleSize = 0,
    ruleFiltration = best_config$ruleFiltration,
    ruleFiltrSupport = best_config$ruleFiltrSupport,
    ruleFiltrAccuracy = best_config$ruleFiltrAccuracy,
    ruleFiltrCoverage = best_config$ruleFiltrCoverage,
    roc = TRUE,
    clroc = levels(selected_positions$class)[1],
    seed = 435)

# --- Genetic (same params, swap reducer + GeneticParam) ---
cat("Running Genetic...\n")
gp_full <- list(
  Modulo      = best_config$jp$Modulo,
  BRT         = best_config$jp$BRT,
  BRTprec     = best_config$jp$BRTprec,
  Precompute  = best_config$jp$Precompute,
  Approximate = best_config$jp$Approximate,
  Fraction    = best_config$jp$Fraction,
  Algorithm   = "Full"
)

ros_genetic <- rosetta(selected_positions,
    reducer = "Genetic",
    GeneticParam = gp_full,
    discreteMethod = best_config$discreteMethod,
    discreteParam = best_config$discreteParam,
    reducerDiscernibility = best_config$reducerDiscernibility,
    underSample = best_config$underSample,
    underSampleNum = best_config$underSampleNum,
    underSampleSize = 0,
    ruleFiltration = best_config$ruleFiltration,
    ruleFiltrSupport = best_config$ruleFiltrSupport,
    ruleFiltrAccuracy = best_config$ruleFiltrAccuracy,
    ruleFiltrCoverage = best_config$ruleFiltrCoverage,
    roc = TRUE,
    clroc = levels(selected_positions$class)[1],
    seed = 435)

# --- Compare ---
cat("\n--- Johnson ---\n")
print(ros_johnson$quality)
cat("Rules:", nrow(ros_johnson$main), "\n")

cat("\n--- Genetic (Simple) ---\n")
print(ros_genetic$quality)
cat("Rules:", nrow(ros_genetic$main), "\n")

cat("\n--- Genetic (Full) ---\n")
print(ros_genetic_full$quality)
cat("Rules:", nrow(ros_genetic_full$main), "\n")
```