---
title: "Project2"
output: html_document
---

# Monte Carlo Feature Selection

Identify the set of relevant features that would be used for classification

```{r}

#install.packages("rJava")
#install.packages("rmcfs")
library(rJava)
library(rmcfs)

# Load data
df <- read.csv("data/clean/positions.csv")

# Drop column ID
positions <- subset(df, select = -id )

# Perform monte carlo feature selection (random reducts)
result <- mcfs(class ~ ., positions, projections = 1500, projectionSize = 0.1, splits = 5, splitSetSize = 500, cutoffPermutations = 6, threadsNumber = 8)

```

```{r}

# Render results
plot(result, type = "distances")

# Fetch results
result2 <- result$RI[1:result$cutoff_value,]

# Keep only relevant features
selected_positions <- positions[, colnames(positions) %in% c(result2$attribute, "class"), drop = FALSE]

# convert from char to factor
selected_positions <- as.data.frame(selected_positions)
selected_positions[] <- lapply(selected_positions, as.factor)

write.csv(selected_positions, "data/clean/selected_positions.csv", row.names = FALSE)
```

# R.ROSETTA Basic Unit Test

Create and train a rule-based model with the selected features 


### Load Library
```{r}
set.seed(435)
library(R.ROSETTA)

# Dimensions
dim(selected_positions)
table(selected_positions$class)
```

### Run Basic Test

```{r}
ros <- rosetta(selected_positions, roc = TRUE, discrete=TRUE, clroc = "Human")

rules <- ros$main
quality <- ros$quality

cat("Johnson - Number of rules:", nrow(rules), "\n")
print(quality)

# print top 12 rules
rlsAut <- viewRules(rules[rules$decision=="Human",])
print(rlsAut[1:12,])

```

### Basic Test Results

```{r}
sig_rules <- rules[rules$pValue < 0.05, ]
cat("Significant rules (p < 0.05):", nrow(sig_rules), "/", nrow(rules), "\n\n")
tabS <- table(sig_rules$decision)
print(tabS)
```

```{r}
for (cls in levels(selected_positions$class)) {
  pct <- tabS[cls] / sum(rules$decision == cls) * 100
    cat("% significant for class", cls, ":", round(pct, 1), "%\n")
}
```

```{r}
# Top rules for each decision class
for (cls in levels(selected_positions$class)) {
    cat("\n--- Top rules for class:", cls, "---\n")
    cls_rules <- rules[rules$decision == cls, ]
    print(viewRules(cls_rules[1:min(10, nrow(cls_rules)), ]))
}
```


```{r}
plotMeanROC(ros)

```

```{r}
gf <- getFeatures(rules, filter = TRUE, filterType = "pvalue", thr = 0.05)
print(gf$features)
```

```{r}
genesAvian  <- gf$features$Avian          # Features occurring in significant avian rules.

genesHuman <- gf$features$Human         # Features occurring in significant human rules.

commongenes <- intersect(genesHuman, genesAvian)
commongenes

uniquegenes <- setdiff(genesHuman, genesAvian)

```

```{r}
rec_rules <- recalculateRules(selected_positions, rules, discrete=TRUE)
```

```{r}
# Top rule for class 0
rules_0 <- rec_rules[rec_rules$decision == "Avian", ]
top_0 <- which.min(rules_0$pValue)
cat("Top rule for class 0 - p =", rules_0$pValue[top_0], "\n")
print(viewRules(rules_0[top_0, ]))

# Top rule for class 1
rules_1 <- rec_rules[rec_rules$decision == "Human", ]
top_1 <- which.min(rules_1$pValue)
cat("\nTop rule for class 1 - p =", rules_1$pValue[top_1], "\n")
print(viewRules(rules_1[top_1, ]))
```

# Clustering Model Rules

Show clustering with heatmap

```{r}

#Hierarchical clustering of model rules
#subsetting the significant rules from the recalculated rules
sig_rec_rules <- rec_rules[rec_rules$pValue < 0.05,]
clusterRules(selected_positions,sig_rec_rules,fontsize = 5)

```

# Visualization

Visualize rules (boxplots, VisuNet, VisuArc)


VisuNet

```{r}
library(RCy3)
library(csVisuNet)
cytoscapePing ()
vis <- visunetcyto(ros$main)
```

VisuArc

```{r}

?VisuArc
```

